---
title: "mars"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(mars)
```

[link](https://github.com/yza608/stat360GroupKDH)  

# Description

The `mars()` function, which is the main function in the **MARS Package**, performs a **Multivariate Adaptive Regression Splines (MARS)** analysis. The **MARS** algorithm builds a regression model by partitioning the predictor variables into segments and fitting a linear regression model to each segment. 


# Usage

> mars(formula, data, control=mars.control())

# Arguments

Argument| Meaning

Formula| interface to specify response and explanatory variables, a variable of the model to be fitted.
Data    | A data frame, list, or environment, containing the variables in the model. If not found in the data, the variables are taken from the environment of *"formula"*
Control | An object of *"mars.control"* with Mmax = 2 and d = 3, by default.


# Details

MARS is a nonparametric regression technique that divides the predictor space into simple regions defined by piecewise linear functions of the input variables. The MARS algorithm  iteratively adds basis functions to the model using a forward stepwise approach. This continues until the specified maximum number of functions or termination criteria is reached. Then, use a backward stepwise approach  to prune the model by removing  basis functions that do not contribute significantly to the model's performance. 
  The mars() function takes as input an expression, data, and a control argument. Returns a list object containing information about the model. This includes calls, equations, responses, basis functions, and the final linear model fit to the data. The control argument of the mars() function is the output of the mars.control() function called within mars(). The `mars.control()` function returns a list of control parameters that can be passed as arguments to the `mars()` function. The standard `mars.control()` contains: 
 
 * Mmax,  maximum number of model terms or basis functions = 2 
 * d,  smoothing parameter to consider or degree of interaction  = 3 
 * trace, a boolean indicating whether the function prints progress messages = FALSE. As Friedman suggested, the user can customize  Mmax and d, but Mmax must be an integer greater than or equal to 2, and d = 3 is optimal. If any validation fails, a warning or error message is printed and the values ​​are reset to  default values. 
  Within mars(), the fwd_stepwise() (based on Algorithm 2) and bwd_stepwise() (based on Algorithm 3) functions are called to perform the forward and backward steps of the MARS algorithm respectively. The LOF() (based on Generalized Cross-Validation or GCV) and h() functions are also called within fwd_stepwise() to compute the lack of non-fitting criteria and use the hinge function respectively to determine the optimal Determine split points. function. The LOF() function is also called inside bwd_stepwise(). The `h()` function has the form $max(0,x-c)$ or $max(0,c-x)$ (these are mirror images).


# Value

`mars()` function returns an object of class *mars”*, including a list that contains at least the following components:

Component| Description
---------|---------
coefficients | A named vector of coefficients
residuals | The residuals, that is response minus fitted value, from the linear regression model.
fitted.values | The fitted mean values from the linear regression model.
rank | The numeric rank of the fitted linear regression model.
df.residual | The residual degrees of freedom.
call | The matched call to the function.
terms | The terms object used.
xlevels | A record of the levels of the factors used in fitting.
y | The response variable used.
B | A matrix with the basis functions used in the model.
Bfuncs | A list of functions used to generate the basis functions.
x_names | The names of the predictor variables.
model | The model frame used.
control | The control parameters used in the model.
qr | The QR decomposition of the model matrix.


# Authors

* **Yiding (Daniel) Zhi** (BSc Data Science)  
* **Kara (My Ngoc Nhu) Nguyen** (BSc Health Sciences, concentration in Life Science)  
* **Harshit Goyal** (BA Economics, concentration in Economic Data Analysis and Minor Statistics)  


# References

* Friedman, J. H. (1991). Multivariate Adaptive Regression Splines. *Annals of Statistics*, 19(1). [link](https://doi.org/10.1214/aos/1176347963)  
* https://archive.ics.uci.edu/ml/machine-learning-databases/00244/
* Wickham, H. (2019). *Advanced R*. CRC Press. [link](https://adv-r.hadley.nz/)  
* Wickham, H. (2015). *R Packages: Organize, Test, Document, and Share Your Code*. “O’Reilly Media, Inc.” [link](https://r-pkgs.org/)  
* https://archive.ics.uci.edu/ml/machine-learning-databases/00294/CCPP.zip
* https://archive.ics.uci.edu/ml/machine-learning-databases/00291/airfoil_self_noise.dat



# See Also

**summary()`** returns a summary table of all independent variable estimate coefficients for the dependent variable. Also provided is the sign (indicating the split direction is left: -1 or right: 1)  for each independent variable and the node for each split. 
 **`anova()`** We first call the function `mars()` to briefly return a summary table of  coefficients for all  variables from the data. `anova.()` also has a similar function to `anova()` to compute the ANOVA table of  one or more linear model fits and return an ANOVA table containing all  degrees of freedom (Df) values. return. Sum of squares (Sum Sq), mean square (Mean Sq), F-statistic (F-value), and P-value (Pr). 
 **predict()`** generates new dependent values ​​based on the dataset.  **`print()`** also calls the `mars()` function and the regression model formula before returning the coefficients for each independent variable. 
  **`plot()`** produces four different residual plots. 
 
 * The **Cumulative Distribution** plot shows the empirical *cumulative distribution function (CDF)* of the data. The empirical CDF is the ratio of values ​​less than or equal to X. It is an increasing step function that exhibits a vertical jump of 1/N for each value of X equal to the observed value. 
 * The **Residuals vs. Fit** plot is a scatterplot of residuals on the y-axis and fitted values ​​(estimated responses) on the x-axis. The chart detects nonlinearities, uneven error variances, and outliers. 
 *  **Normal Q-Q** plots are used to assess how well the distribution of a data set fits a *normal (Gaussian) standard distribution* with mean  0 and  standard deviation  1. 
 * The **Scale-Location** plot plots a function of the residuals that reflects the  variability of the error about the mean. I'm looking for a point cloud that looks random. An increasing relationship means that as the mean increases the variance increases  and vice versa.


